Gate sizing for energy efficient and secure digital designs.

### 2019-05-30 Chapter1/contribution.tex
thesis initial draft

The main contributions of the thesis are as follows:-
\begin{enumerate}
\item We proposed a novel algorithm for the stochastic multi-armed bandit (MAB) problem. Our proposed Efficient UCB Variance method, referred to as EUCBV is an arm elimination algorithm based on UCB-Improved and UCBV strategy which takes into account the empirical variance of the arms and along with aggressive exploration factors eliminate sub-optimal arms. Through a theoretical analysis, we establish that EUCBV achieves a better gap-dependent regret upper bound than UCB-Improved, MOSS, UCB1, and UCBV algorithms. EUCBV enjoys an order optimal gap-independent regret bound same as that of OCUCB and MOSS, and better than UCB-Improved, UCB1 and UCBV. Empirically, in several considered environments EUCBV outperforms most of the state-of-the-art algorithms.

\item We proposed the Augmented-UCB (AugUCB) algorithm for a fixed-budget version of the thresholding bandit problem (TBP), where the objective is to identify a set of arms whose quality is above a threshold. A key feature of AugUCB is that it uses both mean and variance estimates to eliminate arms that have been sufficiently explored. This is the first algorithm to employ such an approach for the considered TBP. Furthermore, in numerical evaluations, we establish in several considered environments that AugUCB outperforms all the algorithms that do not take into consideration the variance of the arms in their action selection strategy.

>>>>

The main contributions of the thesis are as follows:-
\begin{enumerate}
\item We propose the MLTimer alogrithm that uses gate-sizing for reducing the leakage power consumption of a digital design. We propose a smart one-pass tool that can leverage the right optimization technique at the appropriate stage of the flow thereby improving design productivity. A key observation reported in MLTimer is that there exists significant correlation between the timing slacks of gates in the current iteration to the gate replacements in successive iterations. MLTimer leverages this observation to reduce the number of STA runs thereby reducing the overall time taken for optimization.

\item We propose the Karna algorithm which uses gate-sizing for reducing the information leakage via the power side-channel of a digital design. We show that each region in a given design leaks information differently. Thus, it is sufficient to optimize gates in the highly sensitive regions to reduce information leakage. Karna leverages this observation and optimizes gates in these sensitive regions to reduce the power side-channel vulnerability.

### 2019-05-30 Chapter1/feedback.tex
thesis initial draft

\begin{algorithm}[!th]
\caption{An online sequential game}
\label{alg:OSeqGame}
\begin{algorithmic}
\State {\bf Input:} Time horizon $T$, $K$ number of arms with unknown parameters of reward distribution
\State \For{ each timestep $t=1,2,\ldots, T$}
\State The environment chooses a reward vector $F^{env}_{t}= \left[r_{i,t}\sim^{i.i.d} D_{i},\forall i\in\A\right]$.
\State The learner chooses $m$ actions such that $m < K$ following some policy $\pi$, where $\A$ is the set of arms and $|\A|=K$.
\State The learner observes the reward vector $G^{obs}_{t}\subseteq F^{env}_{t}$.
\State \EndFor
\end{algorithmic}
\end{algorithm}

>>>>

% \begin{algorithm}[!th]
% \caption{An online sequential game}
% \label{alg:OSeqGame}
% \begin{algorithmic}
% \State {\bf Input:} Time horizon $T$, $K$ number of arms with unknown parameters of reward distribution
% \State \For{ each timestep $t=1,2,\ldots, T$}
% \State The environment chooses a reward vector $F^{env}_{t}= \left[r_{i,t}\sim^{i.i.d} D_{i},\forall i\in\A\right]$.
% \State The learner chooses $m$ actions such that $m < K$ following some policy $\pi$, where $\A$ is the set of arms and $|\A|=K$.
% \State The learner observes the reward vector $G^{obs}_{t}\subseteq F^{env}_{t}$.
% \State \EndFor
% \end{algorithmic}
% \end{algorithm}

### 2019-05-30 Chapter1/intro.tex
thesis initial draft

In today's world a large number of problems in science and engineering, robotics and 
game playing, resource management, financial portfolio
management, medical treatment design, ad placement, website optimization and packet routing can be modeled as sequential decision-making under uncertainty. Many of these real-world interesting
sequential decision-making problems can be formulated as reinforcement
learning (RL) problems (\cite{bertsekas1996neuro}, \cite{sutton1998reinforcement}). In an RL problem, an agent interacts with a dynamic, stochastic, and unknown environment, with the goal of finding an action-selection strategy or policy that optimizes some long-term performance measure. Every time when the agent interacts with the environment it receives a signal/reward from the environment based on which it modifies its policy. The agent learns to optimize the choice of actions over several time steps which is learned from the sequences of data that it receives from the environment. This is the crux of online sequential learning. This is in contrast to supervised learning methods that deal with labeled data which are independently and identically distributed (i.i.d.) samples from the domain and train some classifier on the entire training dataset to learn the pattern of this distribution to predict future samples (test dataset) with the assumption that it is sampled from the same domain, whereas the RL agent learns from the samples that are collected from the trajectories generated by its sequential interaction with the system. For an RL agent the trajectory consists of a series of sequential interactions whereby it transitions from one state to another following some dynamics intrinsic to the environment while collecting the reward till some stopping condition is reached. This is known as an episode. For a single-step interaction, i.e., when the episode terminates after a single transition, the problem is captured by the multi-armed bandit (MAB) model. Our work will focus on this idea of MAB model.

>>>>

One of the significant moments that has shaped human history was the invention of the computer. What essentially started out as an attempt to replace and replicate simple arithmetic functions now surrounds us on a day-to-day basis taking care of our transport, health, finance and ingratiating itself into our daily lives more and more.

This evolution has largely been made possible due to the invention of the transistor by Bell & Schokley in 1951 which caused computers to go from the size of an airplane hangar to tiny microscopic devices.  The invention of the transistors enabled chip designers to meet the ever increasing user requirements by halving the size of the transistors and thereby doubling the number of transistors that could be packed into the same area. This phenomenon also dubbed as Moore’s law has enabled chip designers meet the exponentially increasing functionality requirements.

In addition to increasing functionality, digital devices are also expected to meet additional constraints which can be enumerated using the tuple {Power, Area and Delay}. Depending on the platform in which it gets deployed, the chip designer might have to optimize the same design to meet  two orthogonal constraints. For example, when building a server, the processor is expected to be fast while the power and area requirements are relaxed but on the other hand if the same processor is to be used on a IoT device, the designer might have to optimise the design for low power, smaller area while the device might not be expected to perform at the same speed as the server. **{ Can we rewrite it with a better example?}

We now discuss the three key objectives that a given digital design is expected to meet. 
\begin{enumerate}
    \item \textbf{Delay} With each passing generation 
\end{enumerate}

%Computers have evolved from being simple calculators to devices that can perform complex operations that are employed in various aspects of our day-to-day lives.  
%This evolution has been

%For the last 70 years, chip designers have been able to meet this expectation by leveraging Moore's law and adding more transistors.

%Today digital devices can be broadly classified into three different classes \begin{enumerate}
 %   \item Embedded devices: These devices are expected to interact with various aspects of a user's environment such as temperature, motion etc. and make decisions on the fly. The emergence of Internet of Things as a popular networking paradigm in the last 3 years has led to an increase in the demand for such devices. These class of devices typically application specific and perform either "send and send" or lightweight computation at the
%\end{enumerate}

%In this evolution, the expectations of the end-user from each generation has been increasing. While initial generations were expected to merely speed-up arithmetic operations, the development of commercial compute devices starting with Intel's 4004 processor has led to these devices perform complex mathematical functions at a faster rate, on a smaller device by using a little energy as possible.

% In today's world, artificial intelligence has proved to be a game-changer in designing agents that interact with an evolving environment and make decisions on the fly. The main goal of artificial intelligence is to design artificial agents that make dynamic decisions in an evolving environment. In pursuit of these, the agent can be thought of as making a series of sequential decisions by interacting with the dynamic environment which provides it with some sort of feedback after every decision which the agent incorporates into its decision-making strategy to formulate the next decision to be made. A large number of problems in science and engineering, robotics and game playing, resource management, financial portfolio management, medical treatment design, ad placement, website optimization and packet routing can be modeled as sequential decision-making under uncertainty. Many of these real-world interesting
% sequential decision-making problems can be formulated as reinforcement learning (RL) problems (\citep{bertsekas1996neuro}, \citep{sutton1998reinforcement}). In an RL problem, an agent interacts with a dynamic, stochastic, and unknown environment, with the goal of finding an action-selection strategy or policy that optimizes some long-term performance measure. Every time when the agent interacts with the environment it receives a signal/reward from the environment based on which it modifies its policy. The agent learns to optimize the choice of actions over several time steps which is learned from the sequences of data that it receives from the environment. This is the crux of online sequential learning.

%     This is in contrast to supervised learning methods that deal with labeled data which are independently and identically distributed (i.i.d.) samples from the considered domain and train some classifier on the entire training dataset to learn the pattern of this distribution to predict the labels of future samples (test dataset) with the assumption that it is sampled from the same domain. In contrast to this, an RL agent learns from the samples that are collected from the trajectories generated by its sequential interaction with the system. For an RL agent, the trajectory consists of a series of sequential interactions whereby it transitions from one state to another following some dynamics intrinsic to the environment while collecting the reward till some stopping condition is reached. This is known as an episode. Here, for an action $i_t$ taken by the agent at the $t$-th timestep, the agent transitions from its current state denoted by $S_{i,t}$ to state $S_{i,t+1}$ and observes the reward $X_{i,t}$. An illustrative image depicting the reinforcement learning scenario is shown in Figure \ref{fig:rl}.

% \begin{figure}[!th]
% \includegraphics[scale=0.45]{Chapter1/img/RL1.png}
% \caption{Reinforcement Learning}
% \label{fig:rl}
% \end{figure}

% %To express an RL problem more formally, we have to define the idea of Markov Decision Process (MDP) which consists of states, actions, transition probabilities, and rewards which in turn helps in deciding the strategy to be followed by the agent.

% %An MDP consists of states

### 2019-05-30 Chapter1/objective.tex
thesis initial draft

The main objectives of the thesis are as follows:-
\begin{enumerate}
\item The first objective of this thesis is to study the area of stochastic multi-armed bandit (SMAB) and how to minimize cumulative regret in this setup. We intend to give strong gap-dependent and gap-independent regret guarantees in the SMAB setting. We also intend to provide the algorithm in the SMAB setting that outperforms the current state-of-the-art algorithms in this setting.

\item The second objective of this thesis is to study the area of thresholding bandit problem (TBP) setting where the goal is to minimize the expected loss at the end of a fixed budget provided as input. We intend to provide strong guarantees with respect to expected loss and also propose the algorithm that does not require any problem complexity as an input. We also intend to provide strong empirical evaluations of the algorithm proposed for the TBP setting.

>>>>

The main objectives of the thesis are as follows:-
\begin{enumerate}
\item The first objective of this thesis is to study the impact of gate-sizing on leakage power minimization. Our problem formulation takes into account the delay constraints of a given design. We also address the issues of scalability and, runtime.

\item The second objective of this thesis is to study the impact of gate-sizing as a possible countermeasure for power side-channel attacks. We propose a gate-sizing algorithm whose objective is to minimize information leakage via the power side-channel. We analyze the effectiveness of this algorithm by employing three different cipher designs. We also discuss the various factors affecting the convergence of the 
%area of thresholding bandit problem (TBP) setting where the goal is to minimize the expected loss at the end of a fixed budget provided as input. We intend to provide strong guarantees with respect to expected loss and also propose the algorithm that does not require any problem complexity as an input. We also intend to provide strong empirical evaluations of the algorithm proposed for the TBP setting.

### 2019-05-30 Chapter1/outline.tex
thesis initial draft

In this chapter, we gave an overview of the various types of bandits available in the literature and also discussed about the main objectives of the thesis and our contributions. In this section, we give a general outline of the thesis that is to follow. In Chapter \ref{chap:SMAB} we give a detailed overview of the stochastic multi-armed bandit model and the latest available algorithms in this setting. In the next Chapter \ref{chap:EUCBV} we introduce our algorithm Efficient UCB Variance (EUCBV) for the stochastic multi-armed bandit model. We give theoretical guarantees on the performance of EUCBV and also show in numerical simulations that it indeed performs very well as compared to the state-of-the-art algorithms. In the subsequent Chapter \ref{chap:tbandit1} we introduce a new variant of pure exploration multi-armed stochastic bandit called the thresholding bandit problem. We analyze the connections between thresholding bandit problem and pure exploration problem and also discuss several existing algorithms in both the settings that are relevant to carefully analyze the thresholding bandit problem. Then in Chapter \ref{chap:tbandit2} we introduce our solution for the thresholding bandit problem, called the Augmented UCB (AugUCB) algorithm. We analyze our algorithm AugUCB and derive theoretical guarantees for it as well as show in numerical experiments that it indeed outperforms several state-of-the-art algorithms in the thresholding bandit setting. Finally, in Chapter \ref{ThesisConc} we conclude by briefly summarizing all the problems covered in the thesis and discussing some future directions in which the stated problems can be further extended.

>>>>

 In this chapter, we gave an overview of the power and security issues in digital designs. We also discussed the gate-sizing technique and its role in solving the above two problems. We now outline the general structure of this thesis. In the next chapter, we discuss MLTimer, a learning based lazy evaluation model that uses gate-sizing for solving the leakage power minimization problem in digital circuits. We discuss Karna, a gate-sizing algorithm for reducing the power side-channels in digital designs in Chapter 3. We compare these two schemes in Chapter 4. We conclude by briefly summarizing the problems covered in this thesis and highlight some interesting future directions that they could be further extended.

% various types of bandits available in the literature and also discussed about the main objectives of the thesis and our contributions. In this section, we give a general outline of the thesis that is to follow. In Chapter \ref{chap:SMAB} we give a detailed overview of the stochastic multi-armed bandit model and the latest available algorithms in this setting. In the next Chapter \ref{chap:EUCBV} we introduce our algorithm Efficient UCB Variance (EUCBV) for the stochastic multi-armed bandit model. We give theoretical guarantees on the performance of EUCBV and also show in numerical simulations that it indeed performs very well as compared to the state-of-the-art algorithms. In the subsequent Chapter \ref{chap:tbandit1} we introduce a new variant of pure exploration multi-armed stochastic bandit called the thresholding bandit problem. We analyze the connections between thresholding bandit problem and pure exploration problem and also discuss several existing algorithms in both the settings that are relevant to carefully analyze the thresholding bandit problem. Then in Chapter \ref{chap:tbandit2} we introduce our solution for the thresholding bandit problem, called the Augmented UCB (AugUCB) algorithm. We analyze our algorithm AugUCB and derive theoretical guarantees for it as well as show in numerical experiments that it indeed outperforms several state-of-the-art algorithms in the thresholding bandit setting. Finally, in Chapter \ref{ThesisConc} we conclude by briefly summarizing all the problems covered in the thesis and discussing some future directions in which the stated problems can be further extended.

### 2019-05-30 Chapter3/introduction.tex
thesis initial draft

>>>>

\section{Introduction}
\label{sec:introduction}

\noindent Dennard's law~\cite{dennard:74} states that the power density would remain constant with technology scaling even with the increase in MOS-device density.  With modern deep nanometer devices, optimization of power consumption has become more complex than optimization for delay or area. This is primarily due to the dominance of leakage power in the sub-100$n$$m$ technology. Techniques like supply voltage scaling and threshold voltage scaling have been used in past to reduce dynamic power consumption while maintaining/improving timing of critical path. However, in sub-100$n$$m$ regime, the exponential increase in sub-threshold leakage due to  threshold voltage scaling has caused leakage power to dominate in microprocessors~\cite{borkar:99,kim:03}. It is also to be noted that leakage power dissipation during idle state does not contribute to any useful computation. In addition, excessive leakage power dissipation can also cause power wastage leading to a  potential {\em thermal runaway}\cite{virat}. This has led to an extensive research on leakage power optimizations at different levels of the VLSI design flow under aggressive timing constraints.

\noindent Gate-sizing and threshold voltage ($V_t$)-sizing are two efficient techniques that are employed at the design stage for reduction of leakage power under timing constraints. In particular, gate-sizing has been very effective in the early and middle stages of the physical synthesis flow.
% * <sristisravan@gmail.com> 2017-06-28T07:45:15.739Z:
% 
% Cite some papers for these statements
% 
% ^.
However, in the post-route stage, gate sizing often necessitates 
incremental placement, which can result in increase of turn-around-time of chip production~\cite{feng:09}. On the other hand, $V_t$ sizing provides enough room for significant optimization in power/timing without any effect on placement. Though, increasing the threshold voltage for a gate (say $g$) by  $V_t$ sizing can result in reducing the leakage power, it can significantly increase the delay. Hence, either gate-sizing or $V_t$ sizing or a mix of both is used to make the optimal choice of each gate depending on the design stage that the chip is present\cite{virat}.

\noindent A standard cell library consists of different versions of the same cell, one for each threshold voltage. However, the number of versions for a given cell is finite and it is limited to the discrete values of the threshold voltages specified by the foundry. Thus, it is imperative to identify cells from the standard cell library that have appropriate gate-size/$V_t$ within the set of available values to minimize the overall leakage power of the circuit without violating the timing constraints.
This optimization problem is called the {\em Timing Constrained Discrete Sizing Problem} (TC-DSP). In this paper, we focus on the TC-DSP for leakage power minimization in digital circuits.

\noindent In addition to the sizing problem, the rapidly shrinking fabrication technology has resulted in an increase in the design optimization effort due to various factors such as multiple process, voltage and temperature corners. This in turn has resulted in an increase in the design effort (design productivity) of chips to match the estimated parameter values at the design phase to what is desirable in the  post fabrication phase. Fast convergence to high quality solutions is thereby required so as to increase design productivity and meet the consumer demand. Practically, the problem of increasing design productivity becomes important during the optimization phase of the VLSI design; as the execution time of this phase spans over hours to days. %For example, reducing optimization time for a given parameter from 2 minutes to 1 second gives more than 2 orders of magnitude speed-up. But from a practical point of view, this reduction is not valuable as the actual reduction is only 119 seconds, which is comparatively very small in the overall design time, that spans several days or months. On the other hand, reducing optimization time from 5 days to 1 hour also gives a $2 \times$ speed-up as explained in the previous case. But in this case, the actual reduction in running time is 4 days and 23 hours, which can significantly improve the design productivity. 
% * <sristisravan@gmail.com> 2017-06-28T07:43:33.360Z:
% 
% 119 seconds
% 
% ^ <slpskp@cse.iitm.ac.in> 2017-06-28T14:34:53.507Z.

\indent ITRS 2011 \cite{itrs:2011} highlights the intensity of research carried out on low power design technology improvements. However, the designers are not able to effectively leverage all the optimization choices due to runtime constraints. For example, consider a design with three cells and a standard cell library with three $V_t$ and ten gate-size choices. If we just consider $V_t$ scaling for power optimization, the search space is $3^3$. Further, if the designer uses gate-sizing, the search space increases to $3^{10}$. Finally, if a mixed sizing technique is employed, the search space increases to $3^{30}$. Thus we can see that, design productivity heavily suffers due to a large exploration space available at each phase of optimization. The authors in  \cite{kahngtalk} list the following three hypothetical steps to improve design productivity without compromising power goals namely:  \begin{itemize}
    \item  \textbf{Optimized back-end:} mixed height library floorplan/placement, ultimate place/route, power/clock distribution;
    \item  \textbf{Modeling/sign-off criteria:} tightened Back End Of Line (BEOL) corners, reduced guardband and Adaptive Voltage Scaling (AVS) aware sign-off; and,
    \item  \textbf{Bespoke,design-specific-flow:} predictive one-pass flow, optimal tool usage.
 \end{itemize} 
 Consider the following pseudocode which represents the template for any iterative greedy gate-sizing optimization algorithm as seen in \cite{hu:12},\cite{hu:13},\cite{mok:12}, and \cite{reiman:13}. The algorithm takes as inputs: the Netlist $C$; containing $N$ gates, that needs to be optimized, a multi-$V_t$ standard cell library, the target frequency $F$ and a function $\alpha(C)$. The function $\alpha(C)$ takes as inputs the circuit C, the timing values for each logic element got by performing a Static Timing Analysis (STA), and other foundry specific parameters for the different logic elements in C, to compute a cost for every gate in C.
\begin{algorithm}
%\algsetup{linenosize=\tiny}
\scriptsize
\LinesNumbered
\title{Greedy Leakage optimization algorithm}
\caption{The template for an iterative greedy leakage optimization algorithm }
\label{alg:naive}
    \KwIn{1) Netlist of the given circuit $C$ containing $N$ gates represented as a Directed Acyclic Graph (DAG). 2) Target frequency $F$. 3) A multi-$V_t$ standard cell library containing multiple cells with different threshold voltages and sizes for every gate; and, 4) Cost function $\alpha(C)$ that computes a cost for every gate in $C$.}
    \KwOut{A leakage optimized netlist running at the target frequency $F$.}
 $N \leftarrow gate\ count$\;

    \textit{Initial Configuration:} Assign to each of the $N$ gates a cell from the standard cell library matching its functionality. The $V_t$ and $size$ of each gate shall vary with different methods employing this template\;
Set iteration count to $1$\;
    Run Static Timing Analysis (STA) and compute $\alpha(C)$\; 
    A $\leftarrow$ list of gates sorted in decreasing order according to cost computed by $\alpha(C)$\;

    \While{A not empty } {
        \textit{Gate replacement:} Consider the first element of A (the gate g with the largest cost) and replace it with its unexplored slower choices (increase $V_t$ or decrease the $size$) so as to reduce the leakage power. Mark the choice as explored for gate g\;
        Estimate the new power and the new delay of the circuit $C$ by performing STA\;
        \If{delay violated} { 
        \textit{Backtrack procedure:} 
            Undo the gate replacement and delete it from A\;
        }
        \Else {
            Increment iteration count\;
            Compute $\alpha(C)$\;
	    If all the choices for gate g, available in the standard cell library, have been explored delete it from A\;
        }
        Sort the gates in A according to decreasing order of cost computed by $\alpha(C)$\;
    }
\end{algorithm}

\noindent It can be observed that the convergence of the above algorithm crucially depends on the initial configuration, the cost function $\alpha(C)$ which is used to determine the candidate gate for replacement and the number of STA calls.  Our proposed leakage optimization technique (\textit{MLTimer}) aims at resolving these bottlenecks. The main contributions of this work are as follows:
\begin{itemize} %check if ieee uses itemize or enumerate??
\item It has been empirically observed that there exists significant correlation between the timing slacks of gates in the current iteration to the gate replacements in the successive iterations. The outcome of this replacement is the fast convergence of the leakage optimization algorithm in lesser number of iterations;
     \item  It can also be seen that a smart one-pass tool that can leverage the right optimization technique at the appropriate stage of the flow can improve design productivity significantly. A novel Support Vector Machine (SVM) based classifier, which provides a good initial design configuration is used to generate a leakage optimal design at the end;
     \item The \textit{MLTimer} algorithm, which leverages the high iterative correlation between timing slacks in the current iteration and the gate replacements in the successive iterations, thereby enabling  {\em lazy timing analysis} to significantly improve the runtime; and, 
\item An {\em Adaptive Window sizing (AWS) scheme for decision making} to perform {\em multiple  replacements} between successive timing updates.

%\item A lazy adaptive heuristic,  that uses the solution provided by the learning algorithm to provide a leakage optimized solution while meeting the delay constraints.
%\item To the best of our knowledge, this is the first work to use learning to solve the leakage optimization problem.
%\item The solutions produced by our algorithm {\em Learntimer} are at least {\em an order of magnitude} faster while retaining the same solution quality as that of iterative greedy heuristics.
%\item We demonstrate the efficiency of our algorithm on the ISPD2012 benchmarks.
\end{itemize}
To the best of our knowledge, this is the first work that employs ML technique to solve the well known leakage optimization problem. We demonstrate the efficiency of our algorithm on both the ISPD2012 benchmarks and a homegrown RISC-V (SHAKTIC) processor that runs linux~\cite{riscv}.

% \begin{figure*}[!ht]
%  \begin{center}
%  \includegraphics[scale=0.45]{fig/predict}
% \captionsetup{singlelinecheck=off}
% \caption [The figure shows the impact of the three hypothetical design steps on solution quality and design time. The proposed steps are Optimized back-end: $(mixed height library floorplan/placement, ultimate place/route, power/clock distribution)$ Modeling/sign-off criteria $(tightened BEOL corners, reduced guardband and AVS aware sign-off)$.\#1: Bespoke,design-specific-flow $(predictive one-pass flow, optimal tool usage)$ ]{The figure shows the impact of the three hypothetical design steps on solution quality (QoR) and design time. The proposed steps are

%  \label{fig:kahng}
%  \end{center}
% \end{figure*}

The rest of the manuscript is organized as follows: Section~\ref{sec:background}  presents the problem along with the literature survey. Section \ref{sec:motivation} describes the need for a learning based leakage optimization technique. Section~\ref{sec:proposed} describes our proposed algorithm while the experimental setup is highlighted in Section~\ref{sec:experiment}. The results are presented in ~\ref{sec:results}. Section~\ref{sec:conclusion} concludes the paper.
% * <sristisravan@gmail.com> 2017-06-28T07:46:32.613Z:
% 
% What does STA mean?
% 
% ^.

### 2019-05-30 abstract.tex
thesis initial draft

This thesis studies the following topics in the area of Reinforcement Learning: Multi-armed bandits in stationary distribution with the goal of cumulative regret minimization and Thresholding bandits in pure exploration setting. The common underlying theme is the study of bandit theory and its application in various types of environments. In the first part of the thesis, we study the classic multi-armed bandit problem with a stationary distribution, one of the first settings studied by the bandit community and which successively gave rise to several new directions in bandit theory. We propose a novel algorithm in this setting and compare both theoretically and empirically its performance against the available algorithms. Our proposed algorithm termed as Efficient-UCB-Variance (EUCBV) is the first arm-elimination algorithm which uses variance estimation to eliminate arms as well as achieve an order optimal regret bound. Empirically, we show that EUCBV outperforms most of the state-of-the-art algorithms in the considered environments. In the next part, we study a specific type of stochastic multi-armed bandit setup called the thresholding bandit problem and discuss its usage, available state-of-the-art algorithms on this setting and our solution to this problem. We propose the Augmented-UCB (AugUCB) algorithm which again uses variance and mean estimation along with arm elimination technique to conduct exploration. We give theoretical guarantees on the expected loss of our algorithm and also analyze its performance against state-of-the-art algorithms in numerical simulations in multiple synthetic environments.

>>>>

Computers have evolved from being glorified calculators that occupied the space of a large building to small pervasive devices. This rapid growth, however, has increased the expectations that the end user places on today's digital devices which, in turn, has increased the number of constraints on the device. This thesis considers two such objectives viz, power and security and proposes design level techniques to address the same.

Today's devices are expected to be extremely performance efficient with both servers and mobile phones expected to run the same workload, e.g., statistical inference algorithms, smaller in size and have a lower energy footprint to enable ubiquity in deployment. This creates challenges for digital designers as these constraints are orthogonal causing designers to sacrifice one or more of these objectives. The exponential growth in the number of EDGE devices has also impacted the device manufacturing time which creates a unique challenge for the digital designer as the design is now expected to meet three orthogonal constraints in a short amount of time. Our proposed MLTimer solution presented in this thesis addresses this problem that uses a novel data-driven algorithm which lets the designer re-use solutions from previously optimized circuits to meet the power constraints for  a given design. We show that the proposed solution outperforms the state-of-the-art solutions by evaluating it on the ISPD2012, ISPD2013 benchmark circuits. We compare the results with a commercial EDA flow by evaluating the proposed solution to optimize a state-of-the-art embedded processor running Linux.

The recent Meltdown and Cambridge Analytica data leaks have sensitized the common man to the importance of a secure software and hardware computing platform. While software vulnerabilities can be fixed by implementing a patch with minimal overheads, hardware vulnerabilities, otherwise known as side-channels, are harder to detect and correct. In this thesis, we consider the problem of power side-channels on embedded devices and propose a gate-sizing based solution to reduce the risk of a digital device through the power side-channel. Our proposed solution, Karna, is the first design level solution. We quantify the effectiveness of Karna along with the area, delay and performance overheads by evaluating the proposed scheme on three different block ciphers.

%This thesis studies the following topics in the area of Reinforcement Learning: classic Multi-armed bandits in stationary distribution with the goal of cumulative regret minimization using variance estimates and Thresholding bandits in pure exploration fixed-budget setting with the goal of instantaneous regret minimization also using variance estimates. The common underlying theme is the study of bandit theory and its application in various types of environments. In the first part of the thesis, we study the classic multi-armed bandit problem with a stationary distribution, one of the first settings studied by the bandit community and which successively gave rise to several new directions in bandit theory. We propose a novel algorithm in this setting and compare both theoretically and empirically its performance against the available algorithms. Our proposed algorithm termed as Efficient-UCB-Variance (EUCBV) is the first arm-elimination algorithm which uses variance estimation to eliminate arms as well as achieve an order optimal regret bound. Empirically, we show that EUCBV outperforms most of the state-of-the-art algorithms in the considered environments. In the next part, we study a specific type of stochastic multi-armed bandit setup called the thresholding bandit problem and discuss its usage, available state-of-the-art algorithms on this setting and our solution to this problem. We propose the Augmented-UCB (AugUCB) algorithm which again uses variance and mean estimation along with arm elimination technique to conduct exploration. We give theoretical guarantees on the expected loss of our algorithm and also analyze its performance against state-of-the-art algorithms in numerical simulations in multiple synthetic environments.
